\chapter*{Acknowledgments}

I would like to thank my advisor, Kirill Levchenko, for pushing me to standards and achievement few attain and for maintaining his support throughout the Ph.D. process.
Thank you to my committee members, Adam Bates, Aaron Schulman, and Gang Wang, for their support and guidance throughout the Ph.D., before, and in the completion of this disseration.

Thank you to Abraham Clements, Gabriela Ciocarlie, and Richard Kennell with CyManII for their support, advice, feedback, particularly in relation to the InteGreat system.
Additionally, I thank my collaborators Anthea Chen, Anushya Iyer, Stephen Checkoway, Stefan Savage, Joshua Mason, YiFei Zhu, Evan Johnson, Mingjia Huo, Stewart Grant, Alex Snoeren, Anil Yelam, and Nishant Bhaskar for their support and efforts throughout our papers together, though the subjects are not all approached in this thesis.
I also thank the University of Illinois, National Science Foundation, and CyManII for funding my research and efforts as a graduate student.

Additional thanks are in order for my labmates Tzu-Bin Yan, Margie Ruffin, Michael Chen, Megan Culler, Jonatas Marques, and many others at the Coordinated Science Laboratory who have provided support and friendship throughout my time at the University of Illinois, including Andrew Miller, Nikita Borisov, Michael Bailey, Deepak Kumar, Zane Ma, Paul Murley, Kyo Kim, and Joshua Reynolds.
I would also like to thank my friends in the Urbana-Champaign community, including Kevin Langowski, Alex Rogers and the Ship of Fools, the Rose Bowl, Waluigi's Mansion, the Spice Rack, and other basements unnamed who supported my harsh noise wall and other musical experiments.

Finally, thank you to Annika Sornson, whose love made me realize not all systems need to be broken: unfortunately, it was too late for the redactions, airplanes, and others.
Thank you as well to her parents for their support and friendship.
Finally, thank you to my parents and step-parents for their love and sacrifices.
Everyone mentioned above is impossible to emulate.
